PyQBench: a Python library for benchmarking gate-based quantum computers

### 2022-12-12 txt/manuscript.tex
drobnostki

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is
positive semidefinite operators having trace equal to one, will be denoted
$\Omega_d$. 
 The subset of $M_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$. 
	An operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a stochastic matrix if $a_{i,j} \ge 0$   and $\sum_{i} a_{i,j} = 1$, whereas an operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a double stochastic matrix if $A$ is a stochastic matrix and $\sum_{j} a_{i,j} = 1$.

@@

\section{Optimal discrimination strategy}
In this Appendix we present proofs of Proposition~\ref{prop-discrim} and Proposition~\ref{prop:optimal-measurement} to indicate the optimal discrimination strategy.
\begin{proof}[Proof of Proposition~\ref{prop-discrim}]
	Let $U = H\diag(1, e^{i \phi}) H^\dagger, \phi \in [0, 
	2\pi)$ be decomposed as 
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvalue $\lambda_1 = 1$, the corresponding 
	eigenvector is 
	of the form $\ket{x_1} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]
	$,
	whereas for  $\lambda_2 = e^{i \phi}$ we have $\ket{x_2} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right] 
	$.
	For Hermitian-preserving maps \cite{watrous} the diamond norm may be expressed as
	\begin{equation}
	\| \Phi  \|_\diamond =  \max_{\rho \in \Omega_d} \| \left( \Id \otimes \sqrt{\rho} \right) \mathcal{J}(\Phi)  \left( \Id \otimes \sqrt{\rho} \right)  \|_1.  \end{equation}
	Hence, we obtain
	\begin{equation}
	\begin{split}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond 
	& =  \max_{\rho \in \Omega_2} \left\| \left( \Id \otimes \sqrt{\rho} \right) 
	\mathcal{J}(\PP_{U} - \PP_{\Id} )  \left( \Id \otimes \sqrt{\rho} \right)  \right\|_1   
	\\ 
	& =  \max_{\rho \in \Omega_2} \left\| \left( \Id \otimes \sqrt{\rho} \right) 
	\sum_{i=0}^{1} \proj{i} \otimes \left( \proj{u_i} - \proj{i} \right)^\top  
	\right\|_1  \\ 
	& = \max_{\rho \in \Omega_2} \left\| \sum_{i=0}^{1} \proj{i} \otimes 
	\sqrt{\rho}  \left( \proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \right\|_1  
	\\ 
	& = \max_{\rho \in \Omega_2} \left\| \sum_{i=0}^{1}\sqrt{\rho}  \left( 
	\proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \right\|_1
	\end{split}
	\end{equation}
	One can prove that for all $\alpha, \beta \ge 0 $, and unit vectors $\ket{x}, 
	\ket{y}$ the following equation holds~\cite{watrous} 
	\begin{equation}
	\| \alpha \proj{x} - \beta\proj{y} \|_1 = \sqrt{(\alpha + \beta)^2 - 4\alpha 
		\beta |\braket{x}{y}|^2}.
	\end{equation}
	By taking $\ket{x} = \frac{\sqrt{\rho} \ket{\bar{u_i}}}{\| \sqrt{\rho} 
		\ket{\bar{u_i}} \|}$ and $ \ket{y} = \frac{\sqrt{\rho} \ket{i}}{\|\sqrt{\rho} 
		\ket{i} \|}$ we have
	\begin{equation}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond  = \max_{\rho \in \Omega_2} 
	\sum_{i=0}^{1} \sqrt{\left( \bra{u_i} \rho \ket{u_i} + \bra{i} \rho \ket{i 
		}\right)^2 - 4 | \bra{u_i} \rho \ket{i} |^2}.
	\end{equation}
	Let us take  $\rho_0 =   \frac{1}{2}  	
	\left(\begin{array}{cc}1&0\\0&1\end{array}\right)  $,   we obtain
	\begin{equation}
	\begin{split}
	||\mathcal{P}_U - \mathcal{P}_{\1}||_\diamond 
	&= \sum_{i=0}^1  
	\sqrt{\left(\bra{u_i}\rho_0\ket{u_i} + \bra{i} \rho_0 \ket{i} \right)^2 - 
		4|\bra{i}\rho_0\ket{u_i}|^2}  \\
	&= \sum_{i=0}^1  \sqrt{ 1 -  \left| \bra{i}  U \ket{i }\right|^2} 
	\\ 
	&=\sum_{i=0}^1  \sqrt{1 -  \left| 1 \cdot \bra{i} \proj{u_1} 
		\ket{i} + e^{i \phi} \cdot\bra{i}  \proj{u_2}\ket{i}\right|^2} \\
	&= \sum_{i=0}^1 
	\sqrt{1 -\left| \frac{1+ e^{i \phi}}{2}\right|^2 } 
	= 2 \sqrt{1 -\left| \frac{1+e^{i \phi}}{2}\right|^2 } \\
	&= |1-e^{i \phi }|.
	\end{split}
	\end{equation}
	Observe that
	\begin{equation}
	\begin{split}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond  
	&=  \| \left( \Id \otimes \sqrt{\rho} \right) \mathcal{J}(\PP_{U} - \PP_{\Id} )  \left( 
	\Id \otimes \sqrt{\rho} \right) \|_1 \\
	&= \left\| ( (\PP_{U} - \PP_\Id) \otimes \Id) \left(  | \sqrt{\rho}^\top 
	\rangle \rangle \langle \langle \sqrt{\rho}^\top | \right) \right\|_1.
	\end{split}
	\end{equation}
	Due to that, we know  
	that $\left| \sqrt{\rho_0}^{\top} \rangle \right\rangle$ is the discriminator of the problem 
	of discrimination between 
	$\PP_{\Id} $ and $\PP_U$ for 
	$ U =  H \diag(1, e^{i \phi}) H^\dagger$ (from definition of diamond norm).      Hence, we obtain that \begin{equation}
	\ket{\psi_{AB}} \coloneqq   | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2} } | 
	\Id_2 \rangle \rangle,
	\end{equation}
	which completes the proof.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:optimal-measurement}]
	From Proposition~\ref{prop-discrim} we obtain the exact form of discriminator given by
	\begin{equation}
	\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
	\rangle \rangle. 
	\end{equation}
	From Holevo-Helstrom theorem  we constrain a conditional measurement $\mu$.  
	Let us define 
	\begin{equation}
	X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
	\otimes \Id_2 \right)(\proj{\psi}).
	\end{equation} 
	From the Hahn-Jordan decomposition, let us note
	\begin{equation}
	X = P - Q
	\end{equation}
	where $P, Q \ge 0 $. 
	Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $\text{im}(P)$ and $\text{im}(Q)$, 
	respectively. Observe, that $P $ and $Q$ are block-diagonal.  Then,  $\Pi_P$ and $\Pi_Q$ have the following forms
	\begin{equation}
	\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
	\end{equation}
	and 
	\begin{equation}
	\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right). 
	\end{equation}
	Hence, we define $V_0$ as
	\begin{equation}
	\begin{cases} V_0 \ket{x_p} = \ket{0} \\ V_0 \ket{x_q} = \ket{1} \end{cases}
	\end{equation}
	and $V_1$ as
	\begin{equation}
	\begin{cases} 
	V_1 \ket{y_p} = \ket{0} \\ 
	V_1 \ket{y_q} = \ket{1}
	\end{cases}. 
	\end{equation}
	For the discrimination task between $\PP_{U}$ and $\PP_{\Id}$, where $U = H 
	\diag(1, e^{i \phi })H$, the explicit form of $V_0$ and $V_1$ is given as 
	follows \todo[inline]{udostepniamy ten plik nb?}(see also \texttt{one-qubit.nb}).
	The unitary $V_0$ has the form
	\begin{equation}
	V_0 = \left(
	\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
	\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
		\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)
	\end{array}
	\right),
	\end{equation}
	whereas the unitary  $V_1$  has the form
	\begin{equation}
	V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
	\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} 
	\right) &  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right),
	\end{equation}
	for all $\phi \in [0,2\pi)$. 
	Finally, we obtain the measurement $\mu$ of the form
	\begin{equation}
	\begin{split}
	\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{0} V_1^\dagger, \\ 
	\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{1} V_1^\dagger.  
	\end{split}
	\end{equation} 
\end{proof}

>>>>

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$M_d$.  
%The subset of $M_d$ consisting of Hermitian matrices of dimension $d$
%will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite
%matrices of dimension $d$ by $\HH_d^+$. 
The set of quantum states, that is
positive semidefinite operators having trace equal to one, will be denoted
$\Omega_d$. 
 The subset of $M_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$. 
	An operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a stochastic matrix if $a_{i,j} \ge 0$   and $\sum_{i} a_{i,j} = 1$, whereas an operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a double stochastic matrix if $A$ is a stochastic matrix and $\sum_{j} a_{i,j} = 1$.

@@

\section{Optimal discrimination strategy}
In this Appendix we present proofs of Proposition~\ref{prop-discrim} and Proposition~\ref{prop:optimal-measurement} to indicate the optimal discrimination strategy.
\begin{proof}[Proof of Proposition~\ref{prop-discrim}]
	Let $U = H\diag(1, e^{i \phi}) H^\dagger, \phi \in [0, 
	2\pi)$ be decomposed as 
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvalue $\lambda_1 = 1$, the corresponding 
	eigenvector is 
	of the form $\ket{x_1} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]
	$,
	whereas for  $\lambda_2 = e^{i \phi}$ we have $\ket{x_2} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right] 
	$.
	For Hermitian-preserving maps \cite{watrous} the diamond norm may be expressed as
	\begin{equation}
	\| \Phi  \|_\diamond =  \max_{\rho \in \Omega_d} \| \left( \Id \otimes \sqrt{\rho} \right) \mathcal{J}(\Phi)  \left( \Id \otimes \sqrt{\rho} \right)  \|_1.  \end{equation}
	Hence, we obtain
	\begin{equation}
	\begin{split}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond 
	& =  \max_{\rho \in \Omega_2} \left\| \left( \Id \otimes \sqrt{\rho} \right) 
	\mathcal{J}(\PP_{U} - \PP_{\Id} )  \left( \Id \otimes \sqrt{\rho} \right)  \right\|_1   
	\\ 
	& =  \max_{\rho \in \Omega_2} \left\| \left( \Id \otimes \sqrt{\rho} \right) 
	\sum_{i=0}^{1} \proj{i} \otimes \left( \proj{u_i} - \proj{i} \right)^\top  
	\right\|_1  \\ 
	& = \max_{\rho \in \Omega_2} \left\| \sum_{i=0}^{1} \proj{i} \otimes 
	\sqrt{\rho}  \left( \proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \right\|_1  
	\\ 
	& = \max_{\rho \in \Omega_2} \left\| \sum_{i=0}^{1}\sqrt{\rho}  \left( 
	\proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \right\|_1
	\end{split}
	\end{equation}
	One can prove that for all $\alpha, \beta \ge 0 $, and unit vectors $\ket{x}, 
	\ket{y}$ the following equation holds~\cite{watrous} 
	\begin{equation}
	\| \alpha \proj{x} - \beta\proj{y} \|_1 = \sqrt{(\alpha + \beta)^2 - 4\alpha 
		\beta |\braket{x}{y}|^2}.
	\end{equation}
	By taking $\ket{x} = \frac{\sqrt{\rho} \ket{\bar{u_i}}}{\| \sqrt{\rho} 
		\ket{\bar{u_i}} \|}$ and $ \ket{y} = \frac{\sqrt{\rho} \ket{i}}{\|\sqrt{\rho} 
		\ket{i} \|}$ we have
	\begin{equation}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond  = \max_{\rho \in \Omega_2} 
	\sum_{i=0}^{1} \sqrt{\left( \bra{u_i} \rho \ket{u_i} + \bra{i} \rho \ket{i 
		}\right)^2 - 4 | \bra{u_i} \rho \ket{i} |^2}.
	\end{equation}
	Let us take  $\rho_0 =   \frac{1}{2}  	
	\left(\begin{array}{cc}1&0\\0&1\end{array}\right)  $,   we obtain
	\begin{equation}
	\begin{split}
	||\mathcal{P}_U - \mathcal{P}_{\1}||_\diamond 
	&= \sum_{i=0}^1  
	\sqrt{\left(\bra{u_i}\rho_0\ket{u_i} + \bra{i} \rho_0 \ket{i} \right)^2 - 
		4|\bra{i}\rho_0\ket{u_i}|^2}  \\
	&= \sum_{i=0}^1  \sqrt{ 1 -  \left| \bra{i}  U \ket{i }\right|^2} 
	\\ 
	&=\sum_{i=0}^1  \sqrt{1 -  \left| 1 \cdot \bra{i} \proj{u_1} 
		\ket{i} + e^{i \phi} \cdot\bra{i}  \proj{u_2}\ket{i}\right|^2} \\
	&= \sum_{i=0}^1 
	\sqrt{1 -\left| \frac{1+ e^{i \phi}}{2}\right|^2 } 
	= 2 \sqrt{1 -\left| \frac{1+e^{i \phi}}{2}\right|^2 } \\
	&= |1-e^{i \phi }|.
	\end{split}
	\end{equation}
	Observe that
	\begin{equation}
	\begin{split}
 \| \left( \Id \otimes \sqrt{\rho} \right) \mathcal{J}(\PP_{U} - \PP_{\Id} )  \left( 
	\Id \otimes \sqrt{\rho} \right) \|_1 = \left\| ( (\PP_{U} - \PP_\Id) \otimes \Id) \left(  | \sqrt{\rho}^\top 
	\rangle \rangle \langle \langle \sqrt{\rho}^\top | \right) \right\|_1.
	\end{split}
	\end{equation}
	Due to that, we know  
	that $\left| \sqrt{\rho_0}^{\top} \rangle \right\rangle$ is the discriminator of the problem 
	of discrimination between 
	$\PP_{\Id} $ and $\PP_U$ for 
	$ U =  H \diag(1, e^{i \phi}) H^\dagger$ (from definition of diamond norm).      Hence, we obtain that \begin{equation}
	\ket{\psi_{0}} \coloneqq   | \sqrt{\rho_0}^\top \rangle \rangle = \frac{1}{\sqrt{2} } | 
	\Id_2 \rangle \rangle,
	\end{equation}
	which completes the proof.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:optimal-measurement}]
	From Proposition~\ref{prop-discrim} we obtain the exact form of discriminator given by
	\begin{equation}
	\ket{\psi_0} = | \sqrt{\rho_0}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
	\rangle \rangle. 
	\end{equation}
	From Holevo-Helstrom theorem  we constrain a conditional measurement $\mu$.  
	Let us define 
	\begin{equation}
	X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi_0}) -  \left( \PP_\Id 
	\otimes \Id_2 \right)(\proj{\psi_0}).
	\end{equation} 
	From the Hahn-Jordan decomposition, let us note
	\begin{equation}
	X = P - Q
	\end{equation}
	where $P, Q \ge 0 $. 
	Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $\text{im}(P)$ and $\text{im}(Q)$, 
	respectively. Observe, that $P $ and $Q$ are block-diagonal.  Then,  $\Pi_P$ and $\Pi_Q$ have the following forms
	\begin{equation}
	\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
	\end{equation}
	and 
	\begin{equation}
	\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right). 
	\end{equation}
	Hence, we define $V_0$ as
	\begin{equation}
	\begin{cases} V_0 \ket{x_p} = \ket{0} \\ V_0 \ket{x_q} = \ket{1} \end{cases}
	\end{equation}
	and $V_1$ as
	\begin{equation}
	\begin{cases} 
	V_1 \ket{y_p} = \ket{0} \\ 
	V_1 \ket{y_q} = \ket{1}
	\end{cases}. 
	\end{equation}
	For the discrimination task between $\PP_{U}$ and $\PP_{\Id}$, where $U = H 
	\diag(1, e^{i \phi })H$, the explicit form of $V_0$ and $V_1$ is given as 
	follows \todo[inline]{udostepniamy ten plik nb?}(see also \texttt{one-qubit.nb}).
	The unitary $V_0$ has the form
	\begin{equation}
	V_0 = \left(
	\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
	\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
		\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)
	\end{array}
	\right),
	\end{equation}
	whereas the unitary  $V_1$  has the form
	\begin{equation}
	V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
	\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} 
	\right) &  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right),
	\end{equation}
	for all $\phi \in [0,2\pi)$. 
	Finally, we obtain the measurement $\mu$ of the form
	\begin{equation}
	\begin{split}
	\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{0} V_1^\dagger, \\ 
	\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{1} V_1^\dagger.  
	\end{split}
	\end{equation} 
\end{proof}
